---
title: "BrCa Regimens"
author: "Alex Paynter"
date: "2023 Apr 17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, include = F,
                      message = F, warning = F)
```

```{r}
library(cli) # prevents an error on version reliance.
library(readr)
library(vctrs)
library(rlang)
library(here)
library(dplyr)
library(tidyr)
library(stringr)
library(magrittr)
library(janitor)
library(glue)
library(genieBPC)
library(ggplot2)
library(sunburstR)

purrr::walk(.x = here("R", dir(here("R"))), .f = source)
```


```{r}
data_list <- readr::read_rds(here("data-raw", "data_list.rds"))
```


## Introduction

This documents shows a few outputs for the breast cancer (BrCa) BioPharma Collaborative (BPC) cohort for AACR project GENIE.  The goal is to facilitate further discussion.

We left the meeting in early April with an interest in looking at medications administered after a progression or recurrence or worsening of disease.  This could take a few forms in the BPC data:

- Regimens started at or after progression by medical oncologist notes.
- Regimens started at or after progression by imaging.
- And/or combinations of medical oncologist note and imaging.
- Regimens started after distant metastases were noted.
- Regimens started after after a period of no cancer, followed by a return of cancer later on (recurrence).

For today we've created some outputs for the first option, regimens started at or after progression by medical oncologist note.  Participants whose death signified the progression are excluded.

```{r}

```


It should be noted that progression was quick in this cohort.  For example


## 