---
title: "GENIE BPC BrCa post-metastasis regimens"
author: "Alex Paynter"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, include = F,
                      message = F, warning = F)
```

```{r}
library(cli) # prevents an error on version reliance.
library(readr)
library(vctrs)
library(rlang)
library(here)
library(dplyr)
library(tidyr)
library(stringr)
library(magrittr)
library(janitor)
library(glue)
library(genieBPC)
library(ggplot2)
library(sunburstR)
library(huxtable)

purrr::walk(.x = here("R", dir(here("R"))), .f = source)
```


```{r}
# data_list <- readr::read_rds(here("data-raw", "data_list.rds"))
load(here("data", "prog_cohorts.Rda"))
```

```{r, calc_stage_iv_no_dmets}
# One calculation needed for insertion into the intro material:
n_stage_iv_no_dmets <- data_list$BrCa_v1.2$ca_dx_index %>% 
  filter(stage_dx_iv %in% "Stage IV") %>% 
  tabyl(ca_dmets_yn) %>%
  # Following includes "Unknown or Not mentioned" as well as missing values in the total.
  filter(!(ca_dmets_yn %in% "Yes")) %>%
  pull(n) %>% 
  sum
```


## Introduction

This documents shows a few outputs for the breast cancer (BrCa) BioPharma Collaborative (BPC) cohort for AACR project GENIE.  

Team members expressed an interest in looking at medications administered on or after identification of distant metastasis of cancer.  This information is encoded differently in PRISSMM for participants who are enrolled at different stages:

- Stage IV - If a patient is diagnosed at stage IV then their time of distant metastasis is the diagnosis date.  An exception is made for participants who are marked stage IV with no distant metastases (currently `r n_stage_iv_no_dmets`).
- Stage 0-III - The time of distant metastasis is directly recorded if observed, and we examine regimens starting on or after this identification.


```{r, num_progressed}
n_brca <- nrow(data_list_cc$BrCa_v1.2$pt_char)
pct_of_cohort <- function(n, d = n_brca) {
  glue("{n} ({formatC(round(n/d*100,1), format = 'f', digits = 1)}%)")
}

n_prog_0d <- dft_dmet_timing %>% filter(tt_d <= 0) %>% nrow %>% pct_of_cohort
n_prog_5y <- dft_dmet_timing %>% filter(tt_y <= 5) %>% nrow %>% pct_of_cohort

```

This criterion produced a large subset, compared to what was expected for `r n_brca` total participants.  For example, `r n_prog_0d` had dmets at diagnosis and `r n_prog_5y` had dmets by 5 years after diagnosis.





## Medication classification

The medications in this sheet are classified starting from a scheme supplied by Jesus Fuentes Antras.  We further iterated to use "Class1" from that sheet, but breaking out several important therapies, such as anti-HER2 targeted regimens, into their own category.  Some notes on the current mapping:

- Medications are broken into the following groups: \{`r dft_drug_map$class_comp %>% unique %>% sort %>% paste(collapse = ", ")`\}.
- These groups are mutually exclusive.
- The groups are **NOT** collectively exhaustive.  The mediations in the original regimen data which are excluded from our outputs here are `r dft_drug_map %>% filter(is.na(class_comp)) %>% pull(agent) %>% unique %>% sort %>% paste(collapse = ", ")`.






```{r, create_sunbursts}
prog_sun_obj_cc <- genieBPC::drug_regimen_sunburst(
  data_list_cc$BrCa_v1.2[c("pt_char", "ca_dx_index", "ca_drugs", "cpt")],
  prog_cohort_cc,
  max_n_regimens = 3
)

set.seed(37) # for color sampling.
# turbo but avoiding the highlighter green.

sun_pal_cc <- make_sun_pal(
  size = nrow(prog_sun_obj_cc$treatment_history)
)

sun_prog_med_onc_cc <- sunburstR::sunburst(
  prog_sun_obj_cc$treatment_history,
  legend = F,
  count = T,
  colors = sun_pal_cc
)


```




## Sunburst plot 

Sunburst plots show regimens started on or after the date of distant metastasis.  The inner ring shows the first regimen, second ring shows the second regimen, and so on.  

*Note:* For participants with multiple index cancers we selected the first one.  This alleviates the potential bias in double counting participants.

```{r, plot-sun-class-one, include = T}
sun_prog_med_onc_cc
```

This figures is limited to the first three regimens patients started on or after a distant metastasis was noted.




```{r, eval = T}
# get a 1-row-per-drug dataset:
dft_drug <- prog_cohort_cc$cohort_ca_drugs %>% 
  # Get only the index cancers:
  left_join(
    select(
      prog_cohort_cc$cohort_ca_dx,
      record_id, ca_seq
    ),
    ., 
    by = c("record_id", "ca_seq")
  ) %>% 
  select(record_id:redcap_ca_index,
         drugs_drug_1:drug_start_end_or_lastadm_int_5) %>%
  mutate(across(
    c(contains("drugs_drug_"), drugs_admin),
    .fn = ~ vec_cast(.x, to = "character"))
  ) %>%
  mutate(
    across(
      .cols = drugs_startdt_int_1:drug_start_end_or_lastadm_int_5,
      .fns = as_double
    )
  )



# split off the character values, we'll do those separately
#   to avoid casting issues.
dft_drug_char <- dft_drug %>%
  select(record_id, 
         regimen_number, 
         ca_seq,
         drugs_drug_1: drugs_drug_5)

dft_drug %<>%
  select(-contains("drugs_drug_")) %>%
  pivot_longer(
    cols = drugs_startdt_int_1:drug_start_end_or_lastadm_int_5,
    names_to = "var",
    values_to = "value"
  ) %>%
  # Because "drug_num" is confusing when we have "drugs_num" in 
  #   the raw data.
  # This is only an id for drug within regimen within person.
  mutate(drug_id = readr::parse_number(var),
         var = stringr::str_replace(var, "_[0-5]", "")) %>%
  pivot_wider(names_from = "var", values_from = "value") %>%
  select(record_id, 
         contains("regimen_number"), 
         drug_id, 
         everything()) 


# Now do it again for the cluster of character columns.
#   All the previous ones were integer/double.
dft_drug_char %<>%
  pivot_longer(
    cols = drugs_drug_1: drugs_drug_5,
    names_to = "var",
    values_to = "value"
  ) %>%
  # Because "drug_num" is confusing when we have "drugs_num" in 
  #   the raw data.
  # This is only an id for drug within regimen within person.
  mutate(drug_id = readr::parse_number(var),
         var = stringr::str_replace(var, "_[0-5]", "")) %>%
  # because there is only one column we don't need to pivot here,
  #  just rename.
  select(record_id, regimen_number, drug_id, drug = value)

dft_drug <-
  left_join(dft_drug, dft_drug_char, 
            by = c("record_id", "regimen_number", "drug_id")) %>%
  relocate(drug, .before = drugs_startdt_int)

# empty rows here have no meaning - it's just regimens with less
#   than 5 drugs which is not suprising or interesting.
dft_drug %<>%
  filter(!is.na(drug))
```


```{r, eval = T}

dfp_top_drugs <- dft_drug %>%
  group_by(record_id, drug) %>%
  summarize(observed = 1, .groups = "drop") %>%
  group_by(drug) %>%
  summarize(n = sum(observed)) %>%
  arrange(desc(n)) %>%
  head(10)

dc_wrap <- function(drug_name) {
  get_drug_compliments(dft_drug, drug_name) %>%
    reshape_drug_compliments()
}

dfp_top_drugs <- dfp_top_drugs %>%
  mutate(df_top_comp = purrr::pmap(
    .l = list(drug_name = drug),
    .f = dc_wrap
  )) %>%
  tidyr::unnest(df_top_comp) %>%
  mutate(drug = str_replace(drug, "\\(.*\\)", ""))
```

## Medications 

The following table shows the drugs taken after a distant metastasis was noted, sorted by the number of participants who took them (n) in any regimen started on or after the dmet.  This only includes drugs associated with the index cancer.

"Compliment 1" shows the drug class most commonly administered along with the drug.  "Compliment 2" shows the second most common.  

```{r, include = T, eval = T}
dfp_top_drugs %>% 
  select(Drug = drug, 
         n = n, 
         `Compliment 1` = first, 
         `Compliment 2` = second) %>%
  huxtable::hux(.) %>%
  theme_compact(.)
  
```






```{r}
dft_bca <- prog_cohort_cc$cohort_ca_dx %>%
  select(record_id, bca_subtype) %>%
  mutate(
    bca_subtype = forcats::fct_na_value_to_level(
      bca_subtype, 
      level = "(Unknown)"
    ),
    bca_subtype = forcats::fct_recode(
      bca_subtype,
      # Stated preference of Pedram's
      `HR+, HER2-` = "HER2-, HR+",
      `HR-, HER2+` = "HER2+, HR-",
      `HR+, HER2+` = "HER2+, HR+"
    ),
    #bca_subtype = forcats::fct_rev(bca_subtype)
  ) %>%
  group_by(bca_subtype) %>%
  mutate(bca_group_ct = n()) %>%
  ungroup()


dft_drug_her_ct <- dft_drug %>%
  select(record_id, drug) %>%
  distinct() %>%
  left_join(., dft_bca, by = "record_id") %>%
  count(bca_subtype, bca_group_ct, drug) %>%
  mutate(
    drug = factor(drug),
    # group these so the targetted therapies are all together:
    drug = forcats::fct_relevel(drug, "antiher2", after = Inf),
    drug = forcats::fct_relevel(drug, "pi3k pathway inhibitor", after = Inf),
    drug = forcats::fct_relevel(drug, "other targetted", after = Inf)
  ) %>%
  mutate(
    prop = n/bca_group_ct,
    str = glue("n={n} ({formatC(prop*100, digits = 1, format = 'f')}%)")
  )

gg_group_her_drug <- ggplot(data = dft_drug_her_ct,
       aes(x = prop,
           fill = bca_subtype,
           y = bca_subtype)) + 
  geom_col() + 
  annotate(geom = 'rect', xmin = 1, xmax = 1.5, ymin = -2, ymax = 10,
           color = "gray80", fill = "gray80") + 
  geom_text(size = 2.5, nudge_x = .05, hjust = 0, 
            aes(label = str)) + 
  facet_wrap(vars(drug),
             ncol = 2,
             dir = "v") + 
  scale_fill_manual(
    name = "Subtype",
    values = c('#4477AA', '#EE6677', '#228833', '#CCBB44', '#66CCEE', '#AA3377')
  ) + 
  scale_x_continuous(
    name = "Proportion of subtype group",
    breaks = c(0,0.5,1),
    labels = paste(c(0, 50, 100), "%"),
    expand = expansion(mult = c(0,0.05), add = c(0,0))
  ) + 
  scale_y_discrete(limits = rev) + 
  coord_cartesian(xlim = c(0,1.2), ylim = c(0.9, 5.1)) + 
  theme_bw() +
  theme(
    axis.title.y = element_blank(),
    legend.position = "none"
  )
```



### HER2/HR status plot (alternate)

Alternative way to view the drug classes and HR/HER status.  Below we show the proporition within each subtype (e.g. within HR+, HER-) who used each class of drugs as a part of a post-dmet regimen associated with the index cancer.  

```{r, plot-gg-her2-hr-drug_c1, include = T}
gg_group_her_drug
```






