---
title: "GENIE BPC BrCa post-metastasis regimens"
author: "Alex Paynter"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, include = F,
                      message = F, warning = F)
```

```{r}
library(cli) # prevents an error on version reliance.
library(readr)
library(vctrs)
library(rlang)
library(here)
library(dplyr)
library(tidyr)
library(stringr)
library(magrittr)
library(janitor)
library(glue)
library(genieBPC)
library(ggplot2)
library(sunburstR)
library(huxtable)

purrr::walk(.x = here("R", dir(here("R"))), .f = source)
```


```{r}
data_list <- readr::read_rds(here("data-raw", "data_list.rds"))



# Selects the first index cancer for participants who have more than one.
# Reasons are:
#   (1) the sunburst plot code won't run correctly with duplicates.
#   (2) there's an obvious strong correlation problem in using two cancers
#       from one person.
data_list$BrCa_v1.2$ca_dx_index <- data_list$BrCa_v1.2$ca_dx_index %>%
  group_by(record_id) %>% 
  slice(1) %>%
  ungroup

```

```{r}

# pt_pfs_m <- get_progressed_timing(
#   ca_dat = data_list$BrCa_v1.2$ca_dx_index,
#   prefix = "pfs_m_adv"
# )
# Update: Now focusing on dmets.
dft_dmet_timing <- get_dmet_timing(
  ca_ind_df = data_list$BrCa_v1.2$ca_dx_index
)
prog_cohort <- filter_dl_by_pt(data_list,
                               dft_dmet_timing)

```


```{r}
# One calculation needed for insertion into the intro material:
n_stage_iv_no_dmets <- data_list$BrCa_v1.2$ca_dx_index %>% 
  filter(stage_dx_iv %in% "Stage IV") %>% 
  tabyl(ca_dmets_yn) %>%
  # Following includes "Unknown or Not mentioned" as well as missing values in the total.
  filter(!(ca_dmets_yn %in% "Yes")) %>%
  pull(n) %>% 
  sum
```


## Introduction

This documents shows a few outputs for the breast cancer (BrCa) BioPharma Collaborative (BPC) cohort for AACR project GENIE.  The goal is to facilitate further discussion.

Team members expressed an interest in looking at medications administered on or after identification of distant metastasis of cancer.  This information is encoded differently in PRISSMM for participants who are enrolled at different stages:


- Stage IV - If a patient is diagnosed at stage IV then their time of distant metastasis is the diagnosis date.  An exception is made for participants who are marked stage IV with no distant metastases (currently `r n_stage_iv_no_dmets`).
- Stage 0-III - The time of distant metastasis is directly recorded if observed, and we examine regimens starting on or after this identification.


```{r}
n_brca <- nrow(data_list$BrCa_v1.2$pt_char)
pct_of_cohort <- function(n, d = n_brca) {
  glue("{n} ({formatC(round(n/d*100,1), format = 'f', digits = 1)}%)")
}

n_prog_0d <- dft_dmet_timing %>% filter(tt_d <= 0) %>% nrow %>% pct_of_cohort
n_prog_5y <- dft_dmet_timing %>% filter(tt_y <= 5) %>% nrow %>% pct_of_cohort

```

This criterion produced a large subset, compared to what was expected for `r n_brca` total participants.  For example, `r n_prog_0d` had dmets at diagnosis and `r n_prog_5y` had dmets by 5 years after diagnosis.


## Sunburst plot

The following sunburst plot shows the first two regimens patients started on or after a progression was noted by medical oncologist.

```{r}
prog_sun_obj <- genieBPC::drug_regimen_sunburst(
  data_list$BrCa_v1.2[c("pt_char", "ca_dx_index", "ca_drugs", "cpt")],
  prog_cohort,
  max_n_regimens = 2
)

set.seed(37) # for color sampling.
# turbo but avoiding the highlighter green.
sun_pal <- sample(
  c(
    viridisLite::turbo(n = nrow(prog_sun_obj$treatment_history)/2,
                       begin = 0.05, end = 0.35),
    viridisLite::turbo(n = nrow(prog_sun_obj$treatment_history)/2,
                       begin = 0.6, end = 0.95)
  )
)
sun_prog_med_onc <- sunburstR::sunburst(
  prog_sun_obj$treatment_history,
  legend = F,
  colors = sun_pal
)
```



```{r, include = T}
sun_prog_med_onc
```

Shows the first two regimens patients started on or after a progression was noted by medical oncologist.








```{r}
# get a 1-row-per-drug dataset:
dft_drug <- prog_cohort$cohort_ca_drugs %>% 
  select(cohort:redcap_ca_index,
         drugs_drug_1:drug_start_end_or_lastadm_int_5) %>%
  mutate(across(
    c(contains("drugs_drug_"), drugs_admin),
    .fn = ~ vec_cast(.x, to = "character"))
  ) %>%
  mutate(
    across(
      .cols = drugs_startdt_int_1:drug_start_end_or_lastadm_int_5,
      .fns = as_double
    )
  )



# split off the character values, we'll do those separately
#   to avoid casting issues.
dft_drug_char <- dft_drug %>%
  select(record_id, regimen_number, 
         drugs_drug_1: drugs_drug_5)

dft_drug %<>%
  select(-contains("drugs_drug_")) %>%
  pivot_longer(
    cols = drugs_startdt_int_1:drug_start_end_or_lastadm_int_5,
    names_to = "var",
    values_to = "value"
  ) %>%
  # Because "drug_num" is confusing when we have "drugs_num" in 
  #   the raw data.
  # This is only an id for drug within regimen within person.
  mutate(drug_id = readr::parse_number(var),
         var = stringr::str_replace(var, "_[0-5]", "")) %>%
  pivot_wider(names_from = "var", values_from = "value") %>%
  select(record_id, contains("regimen_number"), drug_id, everything()) 

# Now do it again for the cluster of character columns.
#   All the previous ones were integer/double.
dft_drug_char %<>%
  pivot_longer(
    cols = drugs_drug_1: drugs_drug_5,
    names_to = "var",
    values_to = "value"
  ) %>%
  # Because "drug_num" is confusing when we have "drugs_num" in 
  #   the raw data.
  # This is only an id for drug within regimen within person.
  mutate(drug_id = readr::parse_number(var),
         var = stringr::str_replace(var, "_[0-5]", "")) %>%
  # because there is only one column we don't need to pivot here,
  #  just rename.
  select(record_id, regimen_number, drug_id, drug = value)

dft_drug <-
  left_join(dft_drug, dft_drug_char, 
            by = c("record_id", "regimen_number", "drug_id")) %>%
  relocate(drug, .before = drugs_startdt_int)

# empty rows here have no meaning - it's just regimens with less
#   than 5 drugs which is not suprising or interesting.
dft_drug %<>%
  filter(!is.na(drug))
```


```{r}


dfp_top_drugs <- dft_drug %>%
  group_by(record_id, drug) %>%
  summarize(observed = 1, .groups = "drop") %>%
  group_by(drug) %>%
  summarize(n = sum(observed)) %>%
  arrange(desc(n)) %>%
  head(10)

dc_wrap <- function(drug_name) {
  get_drug_compliments(dft_drug, drug_name) %>%
    reshape_drug_compliments()
}

dfp_top_drugs <- dfp_top_drugs %>%
  mutate(df_top_comp = purrr::pmap(
    .l = list(drug_name = drug),
    .f = dc_wrap
  )) %>%
  tidyr::unnest(df_top_comp) %>%
  mutate(drug = str_replace(drug, "\\(.*\\)", ""))
```

## Medications

The following table shows the drugs taken after a progression was noted, sorted by the number of participants who took them (n) in any regimen started on or after a progression.  

"Compliment 1" shows the drug most commonly administered in regimens using this row's drug.  "Compliment 2" shows the second most common.  

*Example interpretation using row 1 (may become incorrect with data updates):* Capecitabine was used by more participants after distant metastatses (452) than any other drug.  Regimens containing capecitabine also frequently contained Trastuzumab (10.7% of regimens with Capecitabine contained Trastuzumab) and Lapatinib Ditosylate (8.4%)

```{r, include = T}
dfp_top_drugs %>% 
  select(Drug = drug, 
         n = n, 
         `Compliment 1` = first, 
         `Compliment 2` = second) %>%
  huxtable::hux(.) %>%
  theme_striped(.)
  
```



