---
title: "Survival BrCa"
author: "Alex Paynter"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, include = F,
                      message = F, warning = F)
```

```{r}
library(cli) # prevents an error on version reliance.
library(readr)
library(vctrs)
library(rlang)
library(here)
library(dplyr)
library(tidyr)
library(stringr)
library(magrittr)
library(tibble)
library(janitor)
library(glue)
library(ggplot2)
library(flextable)
library(officer)
library(cfmisc) # https://github.com/alexpaynter/cfmisc/

library(survival)
# library(survminer) # hard pass.
library(ggsurvfit)
library(tranSurv)
library(ggtext)

purrr::walk(
  .x = here("R", dir(here("R"))), 
  .f = source
)
```

```{r,load_data}
# Can later be combined with the other survival analysis script.
dft_surv <- readr::read_rds(
  here("data", "surv_dat.rda")
)

dft_surv %>% 
  mutate(pfs_miss = is.na(tt_pfs_i_and_m_adv_yrs)) %>%
  tabyl(pfs_miss)

```

<style type="text/css">
.main-container {
  max-width: 800px !important;
  margin: auto;
}
</style>

## Terms

- **Delayed cohort entry** or **truncation** both refer to the time of genomic (NGS) sequencing.  This is delayed in our study because participants are only curated at the time of NGS, which is often long after diagnosis of disease.
- **Dependent** truncation refers to when the failure time (e.g., death for overall survival) is associated with the time of cohort entry (NGS).

## Truncation dependence

We expect delayed cohort entry in our study based on Kehl, 2020 (10.1001/jamanetworkopen.2020.6976) and the setup in GENIE BPC.  To test this in our data we will use the estimator of Kendall's conditional tau statistic proposed by Martin and Betensky, 2005 (https://doi.org/10.1198/016214504000001538).  The IPW estimators proposed by Austin and Betensky, 2014 (doi:10.1016/j.csda.2013.11.018.) were found to have lower power for detecting dependent truncation, so we won't use those more complex estimators here.

The following table shows estimates of conditional Kendall's tau.  A nonzero estimate indicates that delayed cohort entry (dependent truncation) was detected.  A negative estimate indicates that long times to sequencing (entry) are associated with short survival times, both relative to diagnosis.

```{r}
# While the Martin and Betensky estimators will be used
#  for confirming dependent cohort entry, we will calculate
#  some other options here to anticipate questions which
#  may arise.

dft_cken <- expand_grid(
  fail = c("tt_os_dx_yrs"),
  method = c("MB", "IPW1", "IPW2")
) %>%
  mutate(
    trunc = "dx_cpt_rep_yrs",
    event = case_when(
      fail %in% "tt_os_dx_yrs" ~ "os_dx_status",
      T ~ NA_character_ 
    )
  ) %>%
  select(trunc, fail, event, method)

dft_cken %<>%
  mutate(
    ckd = purrr::pmap(
      .l = list(t = trunc,
                f = fail,
                e = event,
                m = method),
      .f = (function(t,f,e,m) {
        cken_help(dat = dft_surv,
                  trunc = t,
                  fail = f,
                  event = e,
                  method = m,
                  ci = T)
      })
    )
  ) %>%
  unnest(ckd)

dfp_cken <- dft_cken %>%
  mutate(
    `Cond. Tau` = cfmisc::est_int_str(
      est = est, 
      lower = lcb,
      upper = ucb,
      est_digits = 2
    ),
    `P Value` = pval_nejm(p = p_value)
  ) %>%
  select(`Cond. Tau`, `P Value`, Method = method)
```


```{r, include = T}
dfp_cken %>%
  flextable(.) %>%
  autofit(.) %>%
  color(i = ~Method %in% "MB", color = "black") %>%
  color(i = ~!(Method %in% "MB"), color = 'gray60')
```

## Overall survival



```{r}
surv_obj_os <- with(
  dft_surv,
  Surv(
    time = dx_cpt_rep_yrs,
    time2 = tt_os_dx_yrs,
    event = os_dx_status)
)


surv_obj_os_tm <- with(
  dft_surv,
  trSurvfit(
    trun = dx_cpt_rep_yrs,
    obs = tt_os_dx_yrs,
    delta = os_dx_status,
    tFun = 'linear' # default
  )
)

surv_obj_os_tm # interesting for what they output.

# The following indicates that there is a fit benefit
#   for adding breaks.  However, I'm not clear on how to 
#   fit the model separately and get survival curves out.
# gof_tm_os <- gof(surv_obj_os) 

dft_surv_os_tm <- surv_obj_os_tm$surv %>% 
  as_tibble %>%
  rename_all(tolower)

# Amazing:  confirmed below that this wildly different KM
# fit is risk set adjusted - not the standard KM curve
# assuming no truncation.
```


```{r, plot_surv}

surv_pal <- c('#bb5566', '#004488', '#ddaa33')

vec_surv_subtitle <- glue(
  "<span style = 'color:{surv_pal[1]};'>Transformation model</span> and 
  <span style = 'color:{surv_pal[2]};'>Risk set adjustment  </span>"
)

gg_os <- survfit2(surv_obj_os ~ 1, data = dft_surv) %>%
  ggsurvfit(color = surv_pal[2]) +
  geom_step(data = dft_surv_os_tm, 
            inherit.aes = F,
            aes(x = time, y = trsurv),
            color = surv_pal[1]) + 
  add_risktable(
    risktable_stats = c(
      "n.risk",
      "cum.censor",
      "cum.event"
    )
  ) +
  # too many:
  # add_censor_mark(alpha = 0.1,
  #                 shape = 4,
  #                 size = 2,
  #                 color = surv_pal[3]) +
  # add_quantile(y_value = 0.5,
  #              linetype = 1,
  #              alpha = 0.5,
  #              size = 1,
  #              color = surv_pal[2]) + 
  # these two scale calls are a manual version 
  #   of scale_ggsurvfit()
  # Except that we moved the limits to coord_cartesian
  # to avoid clipping.
  scale_y_continuous(
    expand = c(0,0), 
    label = scales::label_percent()
  ) + 
  scale_x_continuous(
    name = "Years from cancer diagnosis",
    expand = c(0.025,0),
    breaks = seq(0, 100, by = 2.5)
  ) + 
  coord_cartesian(
    xlim = c(0, NA),
    ylim = c(0,1)
  ) + 
  labs(
    title = "Overall survival (from diagnosis)",
    subtitle = vec_surv_subtitle
  ) + 
  theme(
    axis.title.y = element_blank(),
    plot.subtitle = element_markdown()
  )




```


### From diagnosis (whole cohort)

Below we show two methods of estimating overall survival in this cohort:

- **Transformation model** - Using the methods of Chiou, 2019 (https://doi.org/10.1177/09622802188175), this accounts for the correlation between sequencing and death times.  Specifically, we assume there is a transformation which will convert sequencing time and death time into a quasi-independent relationship.  The data are transformed in an attempt to minimize the correlation, and displayed with the parameters that achieve that transformation.
- **Risk set adjustment** - Participants enter the risk set at the time of sequencing.  This model assumes sequencing and death times are independent.  While we believe this assumption is false in our cohort, it is a useful comparator as this is by far the most common method applied to truncated data.

Due to the substantial correlation detected between sequencing and death times, we would expect the risk set adjustment to give incorrect estimates of survival.

```{r, include = T, fig.width = 7, fig.height = 5}
gg_os
```










```{r}

dft_surv_adv <- dft_surv %>%
  filter(stage_dx_iv %in% "Stage IV" | 
           dmets_stage_i_iii %in% 1) %>%
  mutate(
    tt_dmet_dx = if_else(
      stage_dx_iv %in% "Stage IV",
      0,
      dx_to_dmets_yrs),
    tt_os_dmet_yrs = tt_os_dx_yrs - tt_dmet_dx,
    dmet_cpt_rep_yrs = dx_cpt_rep_yrs - tt_dmet_dx,
    # This doesn't need to be done - the survival code fits the same way with or without. 
    # However, it does make me feel better.
    dmet_cpt_rep_yrs = if_else(
      dmet_cpt_rep_yrs < 0, 0, dmet_cpt_rep_yrs
    )
    # tt_pfs_i_and_m_adv_yrs is already from dmets.
  )


surv_obj_os_adv <- with(
  dft_surv_adv,
  Surv(
    time = dmet_cpt_rep_yrs,
    time2 = tt_os_dmet_yrs,
    event = os_dx_status)
)
survfit(surv_obj_os_adv ~ 1)


surv_obj_os_adv_tm <- with(
  dft_surv_adv,
  trSurvfit(
    trun = dmet_cpt_rep_yrs,
    obs = tt_os_dmet_yrs,
    delta = os_dx_status,
    tFun = 'linear' # default
  )
)

surv_obj_os_adv_tm # interesting for what they output.

# The following indicates that there is a fit benefit
#   for adding breaks.  However, I'm not clear on how to 
#   fit the model separately and get survival curves out.
# gof_tm_os <- gof(surv_obj_os) 

dft_surv_os_adv_tm <- surv_obj_os_adv_tm$surv %>% 
  as_tibble %>%
  rename_all(tolower)

# Amazing:  confirmed below that this wildly different KM
# fit is risk set adjusted - not the standard KM curve
# assuming no truncation.
```

```{r}
gg_os_adv <- survfit2(surv_obj_os_adv ~ 1, data = dft_surv) %>%
  ggsurvfit(color = surv_pal[2]) +
  geom_step(data = dft_surv_os_adv_tm, 
            inherit.aes = F,
            aes(x = time, y = trsurv),
            color = surv_pal[1]) + 
  add_risktable(
    risktable_stats = c(
      "n.risk",
      "cum.censor",
      "cum.event"
    )
  ) +
  # too many:
  # add_censor_mark(alpha = 0.1,
  #                 shape = 4,
  #                 size = 2,
  #                 color = surv_pal[3]) +
  # add_quantile(y_value = 0.5,
  #              linetype = 1,
  #              alpha = 0.5,
  #              size = 1,
  #              color = surv_pal[2]) + 
  # these two scale calls are a manual version 
  #   of scale_ggsurvfit()
  # Except that we moved the limits to coord_cartesian
  # to avoid clipping.
  scale_y_continuous(
    expand = c(0,0), 
    label = scales::label_percent()
  ) + 
  scale_x_continuous(
    name = "Years from cancer diagnosis",
    expand = c(0.025,0),
    breaks = seq(0, 100, by = 2.5)
  ) + 
  coord_cartesian(
    xlim = c(0, NA),
    ylim = c(0,1)
  ) + 
  labs(
    title = "Overall survival (from dmet)",
    subtitle = vec_surv_subtitle
  ) + 
  theme(
    axis.title.y = element_blank(),
    plot.subtitle = element_markdown()
  )
```

### From dmet (advanced disease)

The following plot uses the time of advanced disease diagnosis as the index time.  That is, the diagnosis date for Stage IV participants and the noted time of first distant metastasis for everyone else.  This is conditional on survival up to the development of advanced disease.

```{r, include = T, fig.width = 7, fig.height = 5}
gg_os_adv
```

The transformation and risk set adjusted models are similar here because the estimated correlation (Kendall $\tau$) is much weaker in this conditional cohort.  This mirrors the findings of Kehl et. al. (doi: 10.1001/jamanetworkopen.2020.6976), where those with Stage IV disease had little-to-no association between time of sequencing and time of death.


```{r}

dft_surv_adv <- dft_surv %>%
  filter(stage_dx_iv %in% "Stage IV" | 
           dmets_stage_i_iii %in% 1) %>%
  mutate(
    tt_dmet_dx = if_else(
      stage_dx_iv %in% "Stage IV",
      0,
      dx_to_dmets_yrs),
    tt_os_dmet_yrs = tt_os_dx_yrs - tt_dmet_dx,
    dmet_cpt_rep_yrs = dx_cpt_rep_yrs - tt_dmet_dx,
    dmet_cpt_rep_yrs = if_else(
      dmet_cpt_rep_yrs < 0, 0, dmet_cpt_rep_yrs
    )
    # tt_pfs_i_and_m_adv_yrs is already from dmets.
  )


surv_obj_os_adv <- with(
  dft_surv_adv,
  Surv(
    time = dmet_cpt_rep_yrs,
    time2 = tt_os_dmet_yrs,
    event = os_dx_status)
)
survfit(surv_obj_os_adv ~ 1)


surv_obj_os_adv_tm <- with(
  dft_surv_adv,
  trSurvfit(
    trun = dmet_cpt_rep_yrs,
    obs = tt_os_dmet_yrs,
    delta = os_dx_status,
    tFun = 'linear' # default
  )
)

surv_obj_os_adv_tm # interesting for what they output.

# The following indicates that there is a fit benefit
#   for adding breaks.  However, I'm not clear on how to 
#   fit the model separately and get survival curves out.
# gof_tm_os <- gof(surv_obj_os) 

dft_surv_os_adv_tm <- surv_obj_os_adv_tm$surv %>% 
  as_tibble %>%
  rename_all(tolower)
```







## Bivariate time distribution plots

Conceptual work to help with modeling.  The variables here are:

- `tt_pfs_i_and_m_adv_yrs` - progression free survival, only calculated for participants with advanced disease in our data.
- `tt_os_dx_yrs` - overall survival.
- `dx_cpt_rep_yrs` - time of first sequencing report for the participant.


```{r, include = T}
ggplot(dft_surv,
       aes(
         x = tt_pfs_i_and_m_adv_yrs,
         y = tt_os_dx_yrs
       )) + 
  geom_point() +
  coord_equal()

ggplot(dft_surv,
       aes(
         x = dx_cpt_rep_yrs,
         y = tt_pfs_i_and_m_adv_yrs
       )) + 
  geom_point() +
  coord_equal()

ggplot(dft_surv,
       aes(
         x = dx_cpt_rep_yrs,
         y = tt_os_dx_yrs
       )) + 
  geom_point() +
  coord_equal()



```
         
