---
title: "Survival from metastasis"
subtitle: "Version 2 using combined gene features and GOF/LOF"
output: beamer_presentation
date: "Updated 19 Sept 2023"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, include = F,
                      fig.width = 7, fig.height = 5,
                      warning = F, message = F)
```



```{r}
library(fs)
library(purrr)
library(here)
purrr::walk(.x = fs::dir_ls(here("R")), .f = source)
```


```{r}
source_folder_top <- here('data', 'survival', 'v2')
```


```{r}
read_wrap_surv_sum <- function(file) {
  readr::read_rds(
    here(source_folder_top, 'fit_outputs', 'fit_summary', paste0(file, '.rds')
    )
  ) 
}




dft_coef_dmet_os_all <- read_wrap_surv_sum(
     "coef_dmet_os_all"
)
dft_coef_dmet_os_trip_neg <- read_wrap_surv_sum(
     "coef_dmet_os_trip_neg"
)
dft_coef_dmet_os_hr_pos_her2_neg <- read_wrap_surv_sum(
  "coef_dmet_os_hr_pos_her2_neg"
)
dft_coef_dmet_os_her2_pos <- read_wrap_surv_sum(
  "coef_dmet_os_her2_pos"
)


dft_model_compare_dmet_os_all <- read_wrap_surv_sum(
  "model_compare_os_all"
)
dft_model_compare_dmet_os_trip_neg <- read_wrap_surv_sum(
  "model_compare_os_trip_neg"
)
dft_model_compare_dmet_os_hr_pos_her2_neg <- read_wrap_surv_sum(
  "model_compare_os_hr_pos_her2_neg"
)
dft_model_compare_dmet_os_her2_pos <- read_wrap_surv_sum(
  "model_compare_os_her2_pos"
)





# Can replace these once you go back and fix the survival function creator:
dft_dmet_surv_all <- readr::read_rds(
  file = here(source_folder_top, 'prepared_data', 'surv_dmet_all.rds')
)
dft_dmet_surv_trip_neg <- readr::read_rds(
  file = here(source_folder_top, 'prepared_data', 'surv_dmet_trip_neg.rds')
)
dft_dmet_surv_hr_pos_her2_neg <- readr::read_rds(
  file = here(source_folder_top, 'prepared_data', 'surv_dmet_hr_pos_her2_neg.rds')
)
dft_dmet_surv_her2_pos <- readr::read_rds(
  file = here(source_folder_top, 'prepared_data', 'surv_dmet_her2_pos.rds')
)

```

## Model changes

For today:

- Gene features must have a $\geq 2\%$ of cases positive (previously $1\%$).
- Group all alterations for a gene together except in cases where a differential effect is known.  This is done in two ways:

  - EBRR2 is split into mutation and CNA effects (request, previous knowledge).
  - Genes with both gain of function and loss of function alterations are split (the few cases where gain or loss is not annotated are removed).

## Model changes (review from email)

- Plots changed to show log cumulative hazard ratio estimates (previously not log).
- Added comparison with models that do not have confounder adjustments, in other words genes only (+BCA subtype).
  - I've called these models **little** models - the ones with all suspected confounders are **big**.


  


# Results

## Genes with gain and loss   

- The following table shows genes which had both gain and loss annotations from oncoKB (including "likely").  For simplicity I've pared the oncoKB mutation effect terminology down to four categories:

- **Gain** - Likely Gain-of-function and Gain-of-function.
- **Loss** - Likely Loss-of-function and Loss-of-function.
- **Other** - Unknown or inconclusive (or anything else).
- **Switch** - Likely Switch-of-function or Switch-of-function.

I was worried about how to encode switch, but in practice there are too few for it to matter.  If a gene is split into Gain/Loss covariates then we exclude any alterations of Switch or Other type.



```{r calc_diff_effects}
dft_diff_eff <- readr::read_rds(
  here('data', 'genomic', 'gene_gof_lof.rds')
) %>%
  filter(has_gain_and_loss) %>%
  mutate(alt_count = Gain + Other + Loss + Switch) %>%
  arrange(desc(alt_count)) %>%
  select(hugo, Gain, Loss, Switch, Other)

dfp_diff_eff <- dft_diff_eff %>%
  as.data.frame(.) %>%
  mutate(across(.cols = everything(), .fn = as.character)) 
  
```


## Genes with gain and loss (cont.)

Genes with both gain and loss alterations:

```{r show_diff_eff_table, include = T, eval = T}
print(dfp_diff_eff)
```



# Results (survival models)



```{r}
gg_rel_scatter_os_all <- plot_hr_rel_scatter_gof_lof(
  dft_coef_dmet_os_all,
  plot_title = "OS prognostic variables from distant metastasis",
  plot_subtitle = "Reliability is the frequency of identification in a boostrapped regularized LTRC Cox model"
)

gg_rel_scatter_os_trip_neg <- plot_hr_rel_scatter_gof_lof(
  dft_coef_dmet_os_trip_neg,
  plot_title = "OS prognostic variables from distant metastasis (triple negative subgroup)",
  plot_subtitle = "Reliability is the frequency of identification in a boostrapped regularized LTRC Cox model"
)

gg_rel_scatter_os_hr_pos_her2_neg <- plot_hr_rel_scatter_gof_lof(
  dft_coef_dmet_os_hr_pos_her2_neg,
  plot_title = "OS prognostic variables from distant metastasis (HR+/HER2- subgroup)",
  plot_subtitle = "Reliability is the frequency of identification in a boostrapped regularized LTRC Cox model"
)
  
gg_rel_scatter_os_her2_pos <- plot_hr_rel_scatter_gof_lof(
  dft_coef_dmet_os_her2_pos,
    plot_title = "OS prognostic variables from distant metastasis (HER2+ subgroup)",
  plot_subtitle = "Reliability is the frequency of identification in a boostrapped regularized LTRC Cox model"
)
```


```{r}
gg_mod_comp_all <- dft_model_compare_dmet_os_all %>%
  filter(term %in% head(levels(term), 20)) %>%
  plot_lasso_model_comp(.)

gg_mod_comp_trip_neg <- dft_model_compare_dmet_os_trip_neg %>%
  filter(term %in% head(levels(term), 20)) %>%
  plot_lasso_model_comp(.)

gg_mod_comp_her2_pos <- dft_model_compare_dmet_os_her2_pos %>%
  filter(term %in% head(levels(term), 20)) %>%
  plot_lasso_model_comp(.)

gg_mod_comp_hr_pos_her2_neg <- dft_model_compare_dmet_os_hr_pos_her2_neg %>%
  filter(term %in% head(levels(term), 20)) %>%
  plot_lasso_model_comp(.)
```


---

```{r, include = T}
gg_rel_scatter_os_all
```

---

```{r, include = T}
gg_mod_comp_all
```

Plots truncated to show only the top 20.


---

```{r, include = T}
gg_rel_scatter_os_her2_pos
```

---

```{r, include = T}
gg_mod_comp_her2_pos
```






---

```{r, include = T}
gg_rel_scatter_os_trip_neg
```

---

```{r, include = T}
gg_mod_comp_trip_neg
```



---

```{r, include = T}
gg_rel_scatter_os_hr_pos_her2_neg
```

---

```{r, include = T}
gg_mod_comp_hr_pos_her2_neg
```








```{r, include = F, eval = F}
binary_helper <- function(vec, zero = "Wild Type", one = "Altered") {
  case_when(
    vec %in% 0 ~ zero,
    vec %in% 1 ~ one,
    T ~ "error"
  ) %>%
    factor(., levels = c(zero, one))
}

gg_surv_dmet_all_tp53_mut <- dft_dmet_surv_all %>%
  mutate(TP53_mut = binary_helper(TP53_mut)) %>%
  plot_km_dmet_binary_strata(
    dat = .,
    strata = "TP53_mut"
  )

gg_surv_dmet_all_ERBB2_cna <- dft_dmet_surv_all %>%
  mutate(ERBB2_cna = binary_helper(ERBB2_cna)) %>%
  plot_km_dmet_binary_strata(
    dat = .,
    strata = "ERBB2_cna"
  )

gg_surv_dmet_all_stage_iv <- dft_dmet_surv_all %>%
  mutate(stage_dx_iv_num = binary_helper(
    stage_dx_iv_num,
    zero = "Stage I-III at dx",
    one = "Stage IV at dx")
  ) %>%
  plot_km_dmet_binary_strata(
    dat = .,
    strata = "stage_dx_iv_num"
  )
```




```{r, include = F, eval = F}
# left these just to have a template:
gg_surv_dmet_all_tp53_mut
```



```{r, eval = F}
# Tool for me.
dft_n_alt <- bind_rows(
  (
    dft_dmet_surv_all %>%
      select(matches("_mut|_cna|_fus")) %>%
      summarize(
        across(
          .cols = everything(),
          .fns = (function(x) sum(x, na.rm = T))
        )
      ) %>%
      mutate(data = "all")),
  
  (
    dft_dmet_surv_trip_neg %>%
      select(matches("_mut|_cna|_fus")) %>%
      summarize(
        across(
          .cols = everything(),
          .fns = (function(x) sum(x, na.rm = T))
        )
      ) %>%
      mutate(data = "trip_neg")
  ),
  
  (
    dft_dmet_surv_hr_pos_her2_neg %>%
      select(matches("_mut|_cna|_fus")) %>%
      summarize(
        across(
          .cols = everything(),
          .fns = (function(x) sum(x, na.rm = T))
        )
      ) %>%
      mutate(data = "hr_pos_her2_neg")
  ),
  
  (
    dft_dmet_surv_her2_pos %>%
      select(matches("_mut|_cna|_fus")) %>%
      summarize(
        across(
          .cols = everything(),
          .fns = (function(x) sum(x, na.rm = T))
        )
      ) %>%
      mutate(data = "her2_pos")
  )
)

dt_n_alt <- dft_n_alt %>%
  pivot_longer(
    cols = -data,
    names_to = "feature",
    values_to = "n_altered"
  ) %>%
  filter(!is.na(n_altered)) %>%
  DT::datatable(.)

dt_n_alt
```

