---
title: "Breast cancer survival by drug"
subtitle: "Feasibility step - list the most common metastatic regimens by BCA type"
author: "Alex Paynter"
date: "`r format(Sys.Date(), '%d %b %Y')`"
editor_options:
  quarto:
    chunk_output_type: console
  chunk_output_type: console
format:
  html:
    embed-resources: true
    toc: true
    toc-depth: 3
    theme: sandstone 
execute:
  echo: false
  include: false
  warning: false
  message: false
  fig.width: 7
  fig.height: 5
---

```{r}
library(fs); library(purrr); library(here);
purrr::walk(.x = fs::dir_ls(here("R")), .f = source)
```


```{r}

dft_pt <- readr::read_rds(
  here('data', 'clin_data_cohort', 'dft_pt.rds')
)
dft_ca_ind <- readr::read_rds(
  here('data', 'clin_data_cohort', 'dft_ca_ind.rds')
)
dft_reg <- readr::read_rds(
  here('data', 'clin_data_cohort', 'dft_reg.rds')
)

dft_clin_char_dmet <- readr::read_rds(
  here('data', 'survival', 'v2', 'prepared_data', 'clin_char_dmet.rds')
)

dft_clin_char_dmet %<>%
  filter(!is.na(dx_to_dmets_yrs)) %>%
  # removes some of the survival stuff to avoid confusion
  select(
    -c(os_dx_status, tt_os_dx_yrs,
       pfs_i_and_m_adv_status,
       tt_pfs_i_and_m_adv_yrs,
       tt_os_dmet_yrs)
  )
    

dft_reg_met <- left_join(
  dft_clin_char_dmet,
  dft_reg,
  by = c('record_id', 'ca_seq')
)

dft_reg_met %<>%
  # only regmiens started on or after dmet
  mutate(dmet_reg_start_yrs = dx_reg_start_int_yrs - dx_to_dmets_yrs) %>%
  filter(dmet_reg_start_yrs > -0.5/365)

vec_drug_exclusions <- c("Leuprolide Acetate", "Goserlin Acetate")

# This probably isn't the best place to limit on investigational agents, can fix later on.
dft_reg_met %<>%
  filter(
    !str_detect(regimen_drugs, "Investigational"),
    !(regimen_drugs %in% c("Leuprolide Acetate", "Goserlin Acetate"))
  )
                           

dft_bca_nest <- dft_reg_met %>%
  arrange(bca_subtype_f, dmet_reg_start_yrs) %>%
  nest(.by = bca_subtype_f) 

# In an effort to perputaully confuse our readers let's also introduce ER positive.
dft_reg_met_er <- dft_ca_ind %>% 
  select(record_id, ca_seq, ca_bca_er) %>%
  left_join(
    .,
    dft_reg_met,
    by = c('record_id', 'ca_seq')
  ) %>%
  mutate(
    bca_subtype_er_f = case_when(
      is.na(bca_subtype) & ca_bca_er %in% "Positive/elevated"  ~ "ER+, HER2?",
      bca_subtype %in% c("HER2+, HR+", "HER2+, HR-") & ca_bca_er %in% "Positive/elevated"  ~ "ER+, HER2+",
      bca_subtype %in% c("HER2-, HR+", "Triple Negative") & ca_bca_er %in% "Positive/elevated"  ~ "ER+, HER2-",
      T ~ NA_character_ 
    )
  )



dft_bca_nest <- bind_rows(
  mutate(dft_bca_nest, bca_subtype = as.character(bca_subtype_f)),
  (dft_reg_met_er %>%
     arrange(bca_subtype_er_f, dmet_reg_start_yrs) %>%
     nest(.by = bca_subtype_er_f) %>%
     filter(!is.na(bca_subtype_er_f)) %>%
      rename(bca_subtype = bca_subtype_er_f))
) %>%
  select(-bca_subtype_f) 
  
  

dft_bca_nest %<>%
  mutate(
    dat_exposed = purrr::map(
      .x = data,
      .f = \(z) {
        z %>%
          count(record_id, regimen_drugs) %>%
          count(regimen_drugs, sort = T)
      }
    ),
    ft_exposed = purrr::map(
      .x = dat_exposed,
      .f = \(z) {
        z %>%
          slice(1:min(c(n(),10))) %>%
          flextable(.) %>%
          autofit(.)
      }
    )
  )


dft_bca_nest %<>%
  mutate(
    dat_1L = purrr::map(
      .x = data,
      .f = \(z) {
        z %>%
          group_by(record_id) %>%
          arrange(dmet_reg_start_yrs) %>%
          slice(1) %>%
          ungroup() %>%
          count(regimen_drugs, sort = T)
      }
    ),
    ft_1L = purrr::map(
      .x = dat_1L,
      .f = \(z) {
        z %>%
          slice(1:min(c(n(),10))) %>%
          flextable(.) %>%
          autofit(.)
      }
    )
  )

dft_bca_nest %<>%
  mutate(
    dat_cdk = purrr::map(
      .x = data,
      .f = \(z) {
        z %>%
          group_by(record_id) %>%
          filter(str_detect(regimen_drugs, "ciclib")) %>%
          ungroup() %>%
          count(regimen_drugs, sort = T)
      }
    ),
    ft_cdk = purrr::map(
      .x = dat_cdk,
      .f = \(z) {
        z %>%
          slice(1:min(c(n(),10))) %>%
          flextable(.) %>%
          autofit(.)
      }
    )
  )

  
```

## Restrictions

- Excluded any regimen with an investigational agent.
- Excluded these drugs from being considered lines:  `r paste(vec_drug_exclusions, collapse = ", ")`.
  - We can exclude additional drugs if needed get reasonable first line therapy counts below.

## Counts (usual groups)

All of the following counts are for metatstatic regimens.  Fr each BCA subtype we provide two tables:

- **Any met** - the count of people who received a regimen at anytime on or after metastasis.
- **First line** the count of people whose first regimen started after met diagnosis was X.
- **CDK** the count of people exposed to each CDK regimen (anything ending in "ciclib".

Both of these tables are limited to the top 10.

### HR+, HER2+

Any met:

```{r}
#| include: true
dft_bca_nest %>%
  filter(bca_subtype %in% "HR+, HER2+") %>%
  pull(ft_exposed) %>%
  `[[`(.,1)
```

First line:

```{r}
#| include: true
dft_bca_nest %>%
  filter(bca_subtype %in% "HR+, HER2+") %>%
  pull(ft_1L) %>%
  `[[`(.,1)
```

CDK:

```{r}
#| include: true
dft_bca_nest %>%
  filter(bca_subtype %in% "HR+, HER2+") %>%
  pull(ft_cdk) %>%
  `[[`(.,1)
```





### HR+, HER2-

Any met:

```{r}
#| include: true
dft_bca_nest %>%
  filter(bca_subtype %in% "HR+, HER2-") %>%
  pull(ft_exposed) %>%
  `[[`(.,1)
```

First line:

```{r}
#| include: true
dft_bca_nest %>%
  filter(bca_subtype %in% "HR+, HER2-") %>%
  pull(ft_1L) %>%
  `[[`(.,1)
```

CDK:

```{r}
#| include: true
dft_bca_nest %>%
  filter(bca_subtype %in% "HR+, HER2-") %>%
  pull(ft_cdk) %>%
  `[[`(.,1)
```







### HR-, HER2+

Any met:

```{r}
#| include: true
dft_bca_nest %>%
  filter(bca_subtype %in% "HR-, HER2+") %>%
  pull(ft_exposed) %>%
  `[[`(.,1)
```

First line:

```{r}
#| include: true
dft_bca_nest %>%
  filter(bca_subtype %in% "HR-, HER2+") %>%
  pull(ft_1L) %>%
  `[[`(.,1)
```

CDK:

```{r}
#| include: true
dft_bca_nest %>%
  filter(bca_subtype %in% "HR-, HER2+") %>%
  pull(ft_cdk) %>%
  `[[`(.,1)
```




### Triple Negative

Any met:

```{r}
#| include: true
dft_bca_nest %>%
  filter(bca_subtype %in% "Triple Negative") %>%
  pull(ft_exposed) %>%
  `[[`(.,1)
```

First line:

```{r}
#| include: true
dft_bca_nest %>%
  filter(bca_subtype %in% "Triple Negative") %>%
  pull(ft_1L) %>%
  `[[`(.,1)
```

CDK:

```{r}
#| include: true
dft_bca_nest %>%
  filter(bca_subtype %in% "Triple Negative") %>%
  pull(ft_cdk) %>%
  `[[`(.,1)
```



```{r}
#| eval: false

# had a question

# any platinum
dft_bca_nest %>%
  filter(bca_subtype %in% "Triple Negative") %>%
  pull(dat_1L) %>%
  `[[`(.,1) %>%
  filter(
    str_detect(
      tolower(regimen_drugs),
      "carboplatin|cisplatin"
    )
  ) %>%
  pull(n) %>%
  sum


# any plat +/- gem
dft_bca_nest %>%
  filter(bca_subtype %in% "Triple Negative") %>%
  pull(dat_1L) %>%
  `[[`(.,1) %>%
  filter(
    str_detect(
      tolower(regimen_drugs),
      "carboplatin|cisplatin"
    ) &
     str_detect(
      tolower(regimen_drugs),
      "gemcitabine"
    ) 
  ) %>%
  pull(n) %>%
  sum

# any platinum
dft_bca_nest %>%
  filter(bca_subtype %in% "Triple Negative") %>%
  pull(dat_1L) %>%
  `[[`(.,1) %>%
  filter(
    str_detect(
      tolower(regimen_drugs),
      "taxel"
    )
  ) %>%
  pull(n) %>%
  sum

dft_bca_nest %>%
  filter(bca_subtype %in% "Triple Negative") %>%
  pull(dat_1L) %>%
  `[[`(.,1) %>%
  filter(
    str_detect(
      tolower(regimen_drugs),
      "capecita"
    )
  ) %>%
  pull(n) %>%
  sum
    
    
```


### (missing BCA type)

Any met:

```{r}
#| include: true
dft_bca_nest %>%
  filter(is.na(bca_subtype)) %>%
  pull(ft_exposed) %>%
  `[[`(.,1)
```

First line:

```{r}
#| include: true
dft_bca_nest %>%
  filter(is.na(bca_subtype)) %>%
  pull(ft_1L) %>%
  `[[`(.,1)
```

CDK:

```{r}
#| include: true
dft_bca_nest %>%
  filter(is.na(bca_subtype)) %>%
  pull(ft_cdk) %>%
  `[[`(.,1)
```









## Counts (ER-based)

Inferring from our initial discussion, estrogen receptor (ER) status might be more important than HR.  In this section we detail the numbers for the two ER+ groups defined by HER2 status. Obviously these are overlapping with the previous.

### ER+, HER2+


Any met:

```{r}
#| include: true
dft_bca_nest %>%
  filter(bca_subtype %in% "ER+, HER2+") %>%
  pull(ft_exposed) %>%
  `[[`(.,1)
```

First line:

```{r}
#| include: true
dft_bca_nest %>%
  filter(bca_subtype %in% "ER+, HER2+") %>%
  pull(ft_1L) %>%
  `[[`(.,1)
```

CDK:

```{r}
#| include: true
dft_bca_nest %>%
  filter(bca_subtype %in% "ER+, HER2+") %>%
  pull(ft_cdk) %>%
  `[[`(.,1)
```


### ER+, HER2-


Any met:

```{r}
#| include: true
dft_bca_nest %>%
  filter(bca_subtype %in% "ER+, HER2-") %>%
  pull(ft_exposed) %>%
  `[[`(.,1)
```

First line:

```{r}
#| include: true
dft_bca_nest %>%
  filter(bca_subtype %in% "ER+, HER2-") %>%
  pull(ft_1L) %>%
  `[[`(.,1)
```

CDK:

```{r}
#| include: true
dft_bca_nest %>%
  filter(bca_subtype %in% "ER+, HER2-") %>%
  pull(ft_cdk) %>%
  `[[`(.,1)
```





```{r}
#| eval: false
dft_reg_met_er %>% 
  filter(
    !(str_detect(bca_subtype, "HER2+") |
        bca_subtype_er_f %in% "ER+, HER2-") |
      is.na(bca_subtype) | is.na(bca_subtype_er_f)
  ) %>%
  group_by(record_id) %>%
  arrange(dx_reg_start_int) %>%
  slice(1) %>%
  ungroup() %>%
  count(regimen_drugs, sort = T)
```

