---
title: "Survival from metastasis"
output: beamer_presentation
date: "Updated 7 Sept 2023"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, include = F,
                      fig.width = 7, fig.height = 5,
                      warning = F, message = F)
```



```{r}
library(fs)
library(purrr)
library(here)
purrr::walk(.x = fs::dir_ls(here("R")), .f = source)
```




```{r}
read_wrap_surv_sum <- function(file) {
  readr::read_rds(
    here('data', 'survival', 'fit_outputs', 'fit_summary',
         paste0(file, '.rds')
    )
  ) 
}

dft_coef_dmet_os_all <- read_wrap_surv_sum(
     "coef_dmet_os_all"
)
dft_coef_dmet_os_trip_neg <- read_wrap_surv_sum(
     "coef_dmet_os_trip_neg"
)
dft_coef_dmet_os_hr_pos_her2_neg <- read_wrap_surv_sum(
  "coef_dmet_os_hr_pos_her2_neg"
)
dft_coef_dmet_os_her2_pos <- read_wrap_surv_sum(
  "coef_dmet_os_her2_pos"
)
# dft_coef_dmet_os_all_no_conf <- read_wrap_surv_sum(
#   "coef_dmet_os_all_no_conf"
# )
# dft_coef_dmet_os_trip_neg_no_conf <- read_wrap_surv_sum(
#   "coef_dmet_os_trip_neg_no_conf"
# )
# dft_coef_dmet_os_hr_pos_her2_neg_no_conf <- read_wrap_surv_sum(
#   "coef_dmet_os_hr_pos_her2_neg_no_conf"
# )
# dft_coef_dmet_os_her2_pos_no_conf <- read_wrap_surv_sum(
#   "coef_dmet_os_her2_pos_no_conf"
# )


dft_model_compare_dmet_os_all <- read_wrap_surv_sum(
  "model_compare_os_all"
)
dft_model_compare_dmet_os_trip_neg <- read_wrap_surv_sum(
  "model_compare_os_trip_neg"
)
dft_model_compare_dmet_os_hr_pos_her2_neg <- read_wrap_surv_sum(
  "model_compare_os_hr_pos_her2_neg"
)
dft_model_compare_dmet_os_her2_pos <- read_wrap_surv_sum(
  "model_compare_os_her2_pos"
)





# Can replace these once you go back and fix the survival function creator:
dft_dmet_surv_all <- readr::read_rds(
  file = here('data', 'survival', 'prepared_data',
              'surv_dmet_all.rds')
)
dft_dmet_surv_trip_neg <- readr::read_rds(
  file = here('data', 'survival', 'prepared_data',
              'surv_dmet_trip_neg.rds')
)
dft_dmet_surv_hr_pos_her2_neg <- readr::read_rds(
  file = here('data', 'survival', 'prepared_data',
              'surv_dmet_hr_pos_her2_neg.rds')
)
dft_dmet_surv_her2_pos <- readr::read_rds(
  file = here('data', 'survival', 'prepared_data',
              'surv_dmet_her2_pos.rds')
)

```


## Adjustment considerations

At our last meeting we talked about the non-genetic adjustment variables in the model.  The following adjustments were made:

- Age at diagnosis (centered)
- Birth year (centered)
- 0/1 indicator for white primary race.
- 0/1 indicator for hispanic ethnicity.
- An indicator for whether the participant is Stage IV at diagnosis.

Additionally an adjustment for BCA subtype is included for the model of all participants (as opposed to the model stratified by BCA subtype).

We discussed the inclusion of an adjustment for sample type.  I thought about this and I don't think it makes for an appropriate adjustment - the following few slides explain why.

## Adjustment guidelines

VanderWeele (2019) lays out "modified disjunctive cause criterion" for what variables to include in a model.  These are:

1. Any cause of the exposure (except instrumental variables).
2. Any cause of the outcome.
3. Good proxies for unmeasured common causes of the exposure.

I believe age at diagnosis, race and ethnicity are all confounders (affect both exposure and outcome) and sample type is an instrumental variable (affects only the exposure, not related to anything else).  This would mean that adjusting for it would if anything have a detrimental effect on our estimates.


## DAG

```{r, include = T, out.width= '80%'}
knitr::include_graphics(
  here('misc', 'dag.png')
)
```

## Model changes

For today:

- Gene features must have a $\geq 1\%$ of cases positive (previously $0.5\%$).
  - *Note:* Brooke used 2% - I may swap to that to match.
- Plots changed to show log cumulative hazard ratio estimates (previously not log).
- Added comparison with models that do not have confounder adjustments, in other words genes only (+BCA subtype).
  - I've called these models **little** models - the ones with all suspeted confounders are **big**.

## Model changes

Proposed *future* changes: Group all alterations for a gene together except in cases where a differential effect is known.  Two cases where a differential effect is known:

- ERBB2 mutations and CNAs will be split due to a known difference.
- Split any gene which has both alterations with both loss of function and gain of function annotated in our data. 
  - This is in progress (Thanks Brooke!).
  - One example is GATA3, which has some gain of function alterations and some loss of function alterations.  So we will split GATA3 into `GATA3_gof`, `GATA3_lof`.
  


# Results


```{r}



gg_rel_scatter_os_all <- plot_hr_rel_scatter(
  dft_coef_dmet_os_all,
  plot_title = "OS prognostic variables from distant metastasis",
  plot_subtitle = "Reliability is the frequency of identification in a boostrapped regularized LTRC Cox model"
)

gg_rel_scatter_os_trip_neg <- plot_hr_rel_scatter(
  dft_coef_dmet_os_trip_neg,
  plot_title = "OS prognostic variables from distant metastasis (triple negative subgroup)",
  plot_subtitle = "Reliability is the frequency of identification in a boostrapped regularized LTRC Cox model"
)

gg_rel_scatter_os_hr_pos_her2_neg <- plot_hr_rel_scatter(
  dft_coef_dmet_os_hr_pos_her2_neg,
  plot_title = "OS prognostic variables from distant metastasis (HR+/HER2- subgroup)",
  plot_subtitle = "Reliability is the frequency of identification in a boostrapped regularized LTRC Cox model"
)
  
gg_rel_scatter_os_her2_pos <- plot_hr_rel_scatter(
  dft_coef_dmet_os_her2_pos,
    plot_title = "OS prognostic variables from distant metastasis (HER2+ subgroup)",
  plot_subtitle = "Reliability is the frequency of identification in a boostrapped regularized LTRC Cox model"
)
```


```{r}
gg_mod_comp_all <- dft_model_compare_dmet_os_all %>%
  filter(term %in% head(levels(term), 20)) %>%
  plot_lasso_model_comp(.)

gg_mod_comp_trip_neg <- dft_model_compare_dmet_os_trip_neg %>%
  filter(term %in% head(levels(term), 20)) %>%
  plot_lasso_model_comp(.)

gg_mod_comp_her2_pos <- dft_model_compare_dmet_os_her2_pos %>%
  filter(term %in% head(levels(term), 20)) %>%
  plot_lasso_model_comp(.)

gg_mod_comp_hr_pos_her2_neg <- dft_model_compare_dmet_os_hr_pos_her2_neg %>%
  filter(term %in% head(levels(term), 20)) %>%
  plot_lasso_model_comp(.)
```


---

```{r, include = T}
gg_rel_scatter_os_all
```

---

```{r, include = T}
gg_mod_comp_all
```

Plots truncated to show only the top 20.


---

```{r, include = T}
gg_rel_scatter_os_her2_pos
```

---

```{r, include = T}
gg_mod_comp_her2_pos
```






---

```{r, include = T}
gg_rel_scatter_os_trip_neg
```

---

```{r, include = T}
gg_mod_comp_trip_neg
```



---

```{r, include = T}
gg_rel_scatter_os_hr_pos_her2_neg
```

---

```{r, include = T}
gg_mod_comp_hr_pos_her2_neg
```








```{r, include = F, eval = F}
binary_helper <- function(vec, zero = "Wild Type", one = "Altered") {
  case_when(
    vec %in% 0 ~ zero,
    vec %in% 1 ~ one,
    T ~ "error"
  ) %>%
    factor(., levels = c(zero, one))
}

gg_surv_dmet_all_tp53_mut <- dft_dmet_surv_all %>%
  mutate(TP53_mut = binary_helper(TP53_mut)) %>%
  plot_km_dmet_binary_strata(
    dat = .,
    strata = "TP53_mut"
  )

gg_surv_dmet_all_ERBB2_cna <- dft_dmet_surv_all %>%
  mutate(ERBB2_cna = binary_helper(ERBB2_cna)) %>%
  plot_km_dmet_binary_strata(
    dat = .,
    strata = "ERBB2_cna"
  )

gg_surv_dmet_all_stage_iv <- dft_dmet_surv_all %>%
  mutate(stage_dx_iv_num = binary_helper(
    stage_dx_iv_num,
    zero = "Stage I-III at dx",
    one = "Stage IV at dx")
  ) %>%
  plot_km_dmet_binary_strata(
    dat = .,
    strata = "stage_dx_iv_num"
  )
```




```{r, include = F, eval = F}
# left these just to have a template:
gg_surv_dmet_all_tp53_mut
```



```{r, eval = F}
# Tool for me.
dft_n_alt <- bind_rows(
  (
    dft_dmet_surv_all %>%
      select(matches("_mut|_cna|_fus")) %>%
      summarize(
        across(
          .cols = everything(),
          .fns = (function(x) sum(x, na.rm = T))
        )
      ) %>%
      mutate(data = "all")),
  
  (
    dft_dmet_surv_trip_neg %>%
      select(matches("_mut|_cna|_fus")) %>%
      summarize(
        across(
          .cols = everything(),
          .fns = (function(x) sum(x, na.rm = T))
        )
      ) %>%
      mutate(data = "trip_neg")
  ),
  
  (
    dft_dmet_surv_hr_pos_her2_neg %>%
      select(matches("_mut|_cna|_fus")) %>%
      summarize(
        across(
          .cols = everything(),
          .fns = (function(x) sum(x, na.rm = T))
        )
      ) %>%
      mutate(data = "hr_pos_her2_neg")
  ),
  
  (
    dft_dmet_surv_her2_pos %>%
      select(matches("_mut|_cna|_fus")) %>%
      summarize(
        across(
          .cols = everything(),
          .fns = (function(x) sum(x, na.rm = T))
        )
      ) %>%
      mutate(data = "her2_pos")
  )
)

dt_n_alt <- dft_n_alt %>%
  pivot_longer(
    cols = -data,
    names_to = "feature",
    values_to = "n_altered"
  ) %>%
  filter(!is.na(n_altered)) %>%
  DT::datatable(.)

dt_n_alt
```

